---
title: "Mini-Project 2"
author: "Beth Campbell, Lola Roberts and Felicia Villalobos"
date: "March 22, 2018"
output:
  html_document:
    code_folding: hide
---

For this project, we analyzed the FEC data files for the 2011-2012 election cycle. We used information about the candidates, contributions, house elections, and committees to create a visualization about how much money was spent on candidates overall during this cycle, we then used another graphic to compare how many votes the candidates received and whether or not they won the general vote. All of this only applies to the ten candidates with the most money spent on them. These graphics can serve as a visualization for the amount of money spent on candidates and the amount of votes/wins a candidate received in this particular election cycle. 

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
```

```{r, message=FALSE, warning=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```


```{r, message=FALSE, warning=FALSE}
# what we want from candidates
candFilter <- candidates %>%
  select(cand_id, cand_name, cand_party_affiliation, cand_state)

# what we want from contributions
contrFilter <- contributions %>%
  select(cmte_id, name, state, transaction_amt, cand_id)

#joining contributions and candidates
canCon <- candFilter %>%
  full_join(contrFilter, candFilter, by = "cand_id")

# organizing canCon to show what we want
# this is the new data set that has no na's in the cand_name slot 
canCon <- canCon %>%
  filter(!is.na(cand_name),
         !is.na(cmte_id),
         !is.na(name),
         !is.na(state),
         !is.na(transaction_amt),
         transaction_amt > 0) %>%
  group_by(cand_name, cand_state) %>%
  summarize(totalContributionCandidate = sum(transaction_amt)) %>%
  arrange(desc(totalContributionCandidate)) %>%
  head(10)

```

This data may be misleading because we have summed all of the money spent candidates without a distinction of whether the contribution was for or against them. We found the total amount of money spent on each candidate by joining together the candidates dataset and contributions dataset (full_join). Before joining we selected columns from each dataset with relevant information and to speed up the joining process as to not overload the RStudio system. Once we joined them, we filtered out all of the information with NA to make our dataset more clear and easier to understand, from there we grouped by candidate name and state so that we could see the total amount they received. We proceeded to see the candidates that got the largest amount of contributions first by arranging in descending order, our dataset only shows the top ten to make a more readable graph. From this data we can see which candidate got the most contributions and which state that candidate came from. All of the data is for the 2011-2012 election cycle. 

```{r, message=FALSE, warning=FALSE}
# what we want from candidates
candidateSelection <- candidates %>%
  select(cand_id, cand_name, cand_party_affiliation, cand_state)

# what we want from house_elections

houseElection <- house_elections %>%
  select(fec_id, state, candidate_name, general_votes, ge_winner)


houseCandid <- houseElection %>%
  full_join(candidateSelection, houseElection, by = c("fec_id" = "cand_id"))

houseCandContrib <- houseCandid %>%
  inner_join(contrFilter, houseCandid, by = c("fec_id" = "cand_id")) %>%
  filter(!is.na(cand_name),
         !is.na(cmte_id),
         !is.na(name),
         !is.na(transaction_amt),
         !is.na(general_votes),
         transaction_amt > 0) %>%
  mutate(ge_winner = ifelse(ge_winner == "W", 1, 0)) %>%
  group_by(cand_name, cand_state) %>%
  summarize(totalContributionCandidate = sum(transaction_amt),
            general_votes = sum(general_votes),
            ge_winner = sum(ge_winner)) %>%
  mutate(ge_winner = ifelse(ge_winner > 1, "Win", "Lose")) %>%
  arrange(desc(totalContributionCandidate)) %>%
  head(10)

 View(houseCandContrib)
```

```{r, message=FALSE, warning=FALSE}
#graphic showing candidate vs. total contribution
candVs.Money <- ggplot(houseCandContrib, aes(x = cand_name, y = totalContributionCandidate, fill = cand_state)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(size = rel(0.7), angle = 70, hjust = 1)) + 
  xlab("Candidates") + 
  ylab("Total Contribution\n(Millions)") + 
  scale_y_continuous(breaks = c(0e+00, 2e+06, 4e+06, 6e+06), 
                     labels = c(0, 2, 4, 6)) + 
  ggtitle("Candidates vs. Total Contribution")

candVs.Money
```

This data graphic would be a good comparison to the last one in order to see how many votes each candidate got in comparison to the amount of money was spent on them. This data is for the same 10 candidates as in the previous graphic. These two data graphics are meant help us try to find a possible correlation between the amount of money contributed to campaign (either for it or against it) and its effect on the number of general votes. For example, if a candidate has had more money spent on them but, fewer general votes, then maybe we can infer that the money spent was used against their campaign (ex. For a negative ad against a candidate that in turn influenced voters)  

```{r, message=FALSE, warning=FALSE}
candVs.Vote <- ggplot(houseCandContrib, aes(x = cand_name, y = general_votes, fill = factor(ge_winner))) +
geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(size = rel(0.7), angle = 70, hjust = 1)) + 
  xlab("Candidates") + 
  ylab("General Votes\n(10^8)") + 
  scale_y_continuous(breaks = c(0e+00, 1e+08, 2e+08, 3e+08), 
                     labels = c(0, 1, 2, 3)) + 
  ggtitle("Candidates vs. General Votes")

candVs.Vote
```

